\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, fit, calc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{textcomp}
\usepackage{xurl}
\usepackage{pdfpages}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,shapes.misc,fit,backgrounds}
\usepackage{float}

\tikzset{
  block/.style={rectangle, draw, rounded corners, align=center, minimum height=1.2cm, minimum width=3.2cm},
  process/.style={rectangle, draw, align=center, minimum height=1.0cm, minimum width=3.3cm},
  io/.style={trapezium, trapezium left angle=75, trapezium right angle=105, draw, align=center, minimum height=1.0cm, minimum width=3.3cm},
  decision/.style={diamond, draw, aspect=2.0, align=center, inner sep=1pt},
  data/.style={cylinder, draw, shape border rotate=90, aspect=0.25, align=center, minimum height=1.1cm, minimum width=2.0cm},
  line/.style={-{Latex[length=2.2mm,width=2.2mm]}, thick},
  dashedline/.style={-{Latex[length=2mm,width=2mm]}, dashed}
}

\title{Hybrid RSSI--CSI Localization of Hidden IoT Devices in Cluttered Indoor Environments}

\author{
    \IEEEauthorblockN{Mac Dylan Philippe M. Balagtas$^{[1]}$, Niño Angelo U. Balagtas$^{[1]}$, Alec Xavier B. Que Esteves$^{[1]}$, \\ Dr. Luis Gerardo Jr. S. Cañete$^{[2]}$, Dr. Philip Virgil B. Astillo$^{[2]}$}
    \IEEEauthorblockA{\textit{$^{[1]}$Undergraduate Student, Department of Computer Engineering, University of San Carlos, Talamban, Cebu City, Philippines}}
    \IEEEauthorblockA{\textit{$^{[2]}$Faculty Member, Department of Computer Engineering, University of San Carlos, Talamban, Cebu City, Philippines}}
    \IEEEauthorblockA{ 07501508@usc.edu.ph, 24105130@usc.edu.ph, 22102490@usc.edu.ph,}
    \IEEEauthorblockA{ lscanete@usc.edu.ph, pvbastillo@usc.edu.ph}
}

\begin{document}

\maketitle

\begin{abstract}
The proliferation of unauthorized or hidden IoT devices in sensitive indoor environments presents a growing challenge for privacy and security. While Wi-Fi-based Indoor Positioning Systems (IPS) have matured, localizing ``hidden'' stationary nodes in cluttered spaces remains difficult due to severe multipath propagation and signal shadowing. Traditional localization methods rely on Received Signal Strength Indicator (RSSI), which exhibits high variance in indoor settings. This research proposes a robust Real-Time Location System (RTLS) that fuses RSSI with fine-grained Channel State Information (CSI) to enhance localization accuracy for non-cooperative IoT devices. Utilizing a distributed architecture of ESP32 edge nodes and a laptop gateway, the study introduces an adaptive hybrid filter to mitigate indoor RF noise. Experimental results, conducted in a semi-structured classroom environment at the University of San Carlos, demonstrate that the hybrid approach significantly reduces Mean Absolute Error (MAE) compared with baseline RSSI methods, providing a low-cost framework for indoor security and asset tracking.\smallskip
\end{abstract}

\begin{IEEEkeywords}
Indoor Localization, CSI, RSSI, ESP32, Multipath Propagation, Hidden IoT Detection, Physical Layer Security.
\end{IEEEkeywords}

\input{sections/introduction}

\input{sections/rrl.tex}


\input{sections/methodology.tex}

\input{sections/summary.tex}

\bibliographystyle{IEEEtran}
\bibliography{references}

\input{sections/appendix.tex}

\end{document}